---
title: "Vaccines, Quarantine, and the Delta Variant: What We Learned"
author: "Chelsea Rianto"
date: "2025-09-26"
categories: [COVID-19, Quarantine]
---

## Why This Matters

When COVID-19 border quarantine systems were first set up, they worked pretty well at keeping outbreaks out of countries like Australia and New Zealand. But the arrival of the **Delta variant** changed the equation. Delta spread much faster, and vaccines weren’t as effective at stopping transmission. The big question was:  
👉 *Could vaccination alone replace strict quarantine, or did we still need tough border rules?*


## What Data & Models Were Used

The study I reviewed (Zachreson et al., 2022) combined:  

- **Quarantine facility models** tracking travelers, workers, testing, and possible “breach events” (when infection escaped quarantine).  
- **Community outbreak models** that simulated how breaches could spread under different conditions-- vaccine effectiveness, coverage rates, and viral transmissibility.  

The researchers used real-world data on infection rates in travelers, incubation periods, test sensitivity, and Delta’s transmissibility. 


## Key findings

### Figure: Vaccines vs transmissibility  
Here’s a simplified version of one of the study’s results:

```{r}
#| label: fig-outbreak-risk
#| fig-cap: "Outbreak risk depends on both vaccine efficacy and transmissibility. Higher transmissibility (R0) requires more effective vaccines to keep risks under control."
#| echo: false
#| warning: false
#| message: false

library(ggplot2)

df <- expand.grid(R0 = c(3,6,8),
                  VE = seq(0,100,10))

df$risk <- with(df, ifelse(R0==3, 100-VE,
                     ifelse(R0==6, 200-VE*2,
                            300-VE*2.5)))
ggplot(df, aes(VE, risk, color=factor(R0))) +
  geom_line(size=1.2) +
  scale_color_manual(
    values = c("#9ecae1", "#4292c6", "#08519c")  
  ) +
  labs(x="Vaccine Efficacy (%)",
       y="Relative Outbreak Risk",
       color = "R₀"
       ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.background = element_rect(fill = "#fdfdfd", color = NA), 
    plot.background = element_rect(fill = "#fdfdfd", color = NA),
    axis.title.x = element_text(face = "bold"),   
    axis.title.y = element_text(face = "bold"),   
    legend.title  = element_text(face = "bold")
  )
```
We can see that for mild strains (*R₀ ≈ 3*), vaccines weren’t strictly necessary.  
But for **Delta (*R₀ ≈ 6*)**, vaccines needed to be **at least 60% effective** to keep outbreak risks at baseline.  
For even more transmissible strains (*R₀ ≈ 8*), they had to be **above 70%**.

## Community vaccination matters

The study shows that **population vaccination rates** change outcomes meaningfully.
See @tbl-vax-coverage.

```{r}
#| label: tbl-vax-coverage
#| tbl-cap: "Impact of Vaccination Coverage on Time to Outbreak"
#| echo: false
#| warning: false
#| message: false

library(dplyr)
library(gt)

tibble(
  Coverage = factor(c("60%", "70%", "80%", "90%"),
                    levels = c("60%", "70%", "80%", "90%")),
  `Time to Outbreak` = c("Short (weeks)", "Moderate", "More than doubled", "Much longer")
) %>%
  gt() %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(everything())
  )
```
When coverage reached **80% or higher**, the time until an outbreak *more than doubled*.
If coverage was lower (below 60%), outbreaks tended to occur quickly no matter what. See @tbl-vax-coverage.

## Conclusion
The results are clear:

**🟢 Vaccination helps a lot, but quarantine remains necessary for highly transmissible variants.**

Even if all travelers and staff were vaccinated, breaches still occurred. Without strong quarantine, outbreaks followed quickly — and if the community wasn’t highly vaccinated, spread was much faster.
Instead of treating *vaccines vs. quarantine* as a trade-off, policymakers should treat them as *complementary strategies*: vaccines reduce infection risk, while quarantine reduces the chance of importation.
These findings align with Zachreson et al. (2022), who concluded that vaccine coverage alone could not fully replace strict border quarantine during the Delta wave.

## References
Zachreson, C., Shearer, F. M., Price, D. J., Lydeamore, M. J., McVernon, J., McCaw, J., & Geard, N. (2022). COVID-19 in low-tolerance border quarantine systems: Impact of the Delta variant of SARS-CoV-2. Science advances, 8(14), eabm3624. [https://doi.org/10.1126/sciadv.abm3624](https://doi.org/10.1126/sciadv.abm3624).
